The impact of information about tobacco-related reproductive vs. general health risks on South Indian women's tobacco use decisions

### 2020-11-06 paper.Rmd
Final submission tweaks

**Exploratory analyses.** We used a logistic elasticnet regression model to explore baseline tobacco user status as a function of age, pregnancy status, tobacco use by family or friends, education, income, marital status, arranged marriage, number of children, hours of domestic work, hours of non-domestic work, and number of harms mentioned at baseline. Positive predictors of tobacco user status were age, income, and number of children, whereas being married or pregnant were negative predictors. For details, see the SI and Figure S\@ref(fig:user).

@@

Table 2. Mixed effects regression models of the effect of the intervention on different outcomes, as specified in the table headings. Values are estimated coefficients (standard errors). All models included presentation type (the intervention), control for pregnancy status, and included a random intercept for presentation group. Base levels: Trimester 1 and General harms presentation (GHP). **Tobacco self-report**: Negative binomial mixed-effects regression model of self-reported tobacco use frequency at followup, controlling for self-reported baseline frequency. **Cotinine 1**: Linear mixed-effects regression model of followup cotinine concentrations, controlling for baseline cotinine concentrations. **Cotinine 2**: Same as Cotinine 1, except includes an interaction term between baseline cotinine and presentation condition. See Figure \@ref(fig:m5post). **Number of harms**: Poisson mixed effects regression model of the total number of harms mentioned at followup, controlling for number of harms mentioned at baseline. Harms distinguished by general vs. reproductive. This model has double the number of observations because there were two separate counts of harm per participant (reproductive and general); the random effect was thus participant nested within presentation group. See Figure \@ref(fig:harmnum). **Share presentation**: Logistic regression model of sharing information from the presentation as a function of presentation type and the number of harms mentioned at followup. Coefficients are log odds. See Figure S\@ref(fig:sharinglogistic). For more model statistics, see Table SY.

@@

Finally, because individuals who exhibited the greatest increases in number of free-listed harms at followup were arguably most heavily influenced by the presentations, we fit an exploratory model of followup cotinine as a function of the changes in number of general harms and reproductive harms mentioned at followup, relative to baseline. Increases in free-listed reproductive harms, but not general harms, were significantly negatively associated with followup cotinine. See Table SX and Figure S\@ref(fig:cotharmfig).

@@

Of the `r nrow(followup)` participants at followup, `r sum(followup$presentation_share2 == 'yes')` reported that they shared information from the presentation with someone else, e.g., a relative, friend, or neighbor. We fit a mixed effects logistic regression model of sharing information as a function of the number of harms mentioned at followup, presentation condition (RHP or GHP), and trimester. Participants in the RHP condition were slightly but not significantly more likely than those in GHP condition to share information from the presentation. The more harms a participant free-listed at followup the more likely she was to share the information with others. There was no significant effect of trimester or non-pregnant status. See Table 2, Share presentation model, and Figure S\@ref(fig:sharinglogistic).

@@

Community-based research has found that peer-to-peer transmission of tobacco harms is an effective strategy in reducing use because peers are less intimidating than physicians [@secker2000; @williams2011]. Most participants reported that they shared information from the presentations with others, although women in the RHP condition were not more likely to share than women in the GHP condition, contrary to predictions (see Figure S\@ref(fig:sharinglogistic) and Table 2, Share presentation model). This finding nonetheless indicates that sharing in this community is an effective strategy for communicating public health knowledge, and future research should consider using peer-to-peer transmission of tobacco harms to promote cessation in this community.

One exploratory result was particularly noteworthy. The women who were arguably most heavily influenced by the presentations where those who exhibited the greatest increases in the number of free-listed harms at followup relative to baseline. In support, an exploratory analysis found that an increase in the freelisted number of reproductive harms, but not general harms, was a significant negative predictor of cotinine. See Table SX and Figure S\@ref(fig:cotharmfig).

A second exploratory analysis found that probable use of tobacco was positively associated with older age, income, and number of children, and a few other variables, and negatively associated with being married and pregnant. The positive effect of age is consistent with the higher prevalence of tobacco use seen globally in older women [@ng2014], and this pattern is consistent with the fetal protection hypothesis because the older women in our study were approaching the end of their reproductive careers [and therefore would be less likely to be, or become, pregnant; @hagen2013; @hagen2016]. The positive effect of income with use is consistent with gender inequality models [@hitchman2011], but the positive effect of number of children is not. The negative effect of married status with tobacco use is also seen in other populations [@ramsey2019], and the negative effect of pregnancy status is consistent with the fetal protection model. See Figure S\@ref(fig:user).

@@

$$
penalty = (1-\alpha)/2||\beta||_2^2+\alpha||\beta||_1.
$$

@@

Individuals who exhibited the greatest increases in number of free-listed harms at followup were arguably most heavily influenced by the presentations. We therefore fit an exploratory mixed effects linear regression model of log10 followup cotinine as a function of the changes in numbers of free-listed general harms and reproductive harms, controlling for log10 baseline cotinine and Presentation. Increases in the numbers of both types of free-listed harms were negatively associated with followup cotinine, but more so for reproductive harms. See Table SY.

@@

Table SY: Linear mixed effects model of log10 followup cotinine concentration as a function of the changes in numbers of general and reproductive harms free-listed at followup compared to baseline, controlling for log10 baseline cotinine and presentation condition. Values are estimated coefficients (standard errors).

```{r cotharmfig, fig.width=10, fig.cap="Effects plot for a regression model of log10 cotinine concentration versus the change in number of (A) general harms and (B) reproductive harms mentioned at followup compared to baseline (see Table SY)."}
p_m_cot_deltargharmsGH + p_m_cot_deltargharmsRH + plot_annotation(tag_levels='A')
```

@@

This table displays the coefficents of the same regression models as Table X, but adds z-values, full p-values, and 95% CI's.

@@

>>>>

**Exploratory analyses.** We used a logistic elasticnet regression model to explore baseline tobacco user status as a function of age, pregnancy status, tobacco use by family or friends, education, income, marital status, arranged marriage, number of children, hours of domestic work, hours of non-domestic work, and number of harms mentioned at baseline. Positive predictors of tobacco user status were age, income, and number of children, whereas being married or pregnant were negative predictors. For details, see the SI and Figure S4.

@@

Table 2. Mixed effects regression models of the effect of the intervention on different outcomes, as specified in the table headings. Values are estimated coefficients (standard errors). All models included presentation type (the intervention), control for pregnancy status, and included a random intercept for presentation group. Base levels: Trimester 1 and General harms presentation (GHP). **Tobacco self-report**: Negative binomial mixed-effects regression model of self-reported tobacco use frequency at followup, controlling for self-reported baseline frequency. **Cotinine 1**: Linear mixed-effects regression model of followup cotinine concentrations, controlling for baseline cotinine concentrations. **Cotinine 2**: Same as Cotinine 1, except includes an interaction term between baseline cotinine and presentation condition. See Figure \@ref(fig:m5post). **Number of harms**: Poisson mixed effects regression model of the total number of harms mentioned at followup, controlling for number of harms mentioned at baseline. Harms distinguished by general vs. reproductive. This model has double the number of observations because there were two separate counts of harm per participant (reproductive and general); the random effect was thus participant nested within presentation group. See Figure \@ref(fig:harmnum). **Share presentation**: Logistic regression model of sharing information from the presentation as a function of presentation type and the number of harms mentioned at followup. Coefficients are log odds. See Figure S6. For more model statistics, see Table S2.

@@

Finally, because individuals who exhibited the greatest increases in number of free-listed harms at followup were arguably most heavily influenced by the presentations, we fit an exploratory model of followup cotinine as a function of the changes in number of general harms and reproductive harms mentioned at followup, relative to baseline. Increases in free-listed reproductive harms, but not general harms, were significantly negatively associated with followup cotinine. See Table S1 and Figure S5.

@@

Of the `r nrow(followup)` participants at followup, `r sum(followup$presentation_share2 == 'yes')` reported that they shared information from the presentation with someone else, e.g., a relative, friend, or neighbor. We fit a mixed effects logistic regression model of sharing information as a function of the number of harms mentioned at followup, presentation condition (RHP or GHP), and trimester. Participants in the RHP condition were slightly but not significantly more likely than those in GHP condition to share information from the presentation. The more harms a participant free-listed at followup the more likely she was to share the information with others. There was no significant effect of trimester or non-pregnant status. See Table 2, Share presentation model, and Figure S6.

@@

Community-based research has found that peer-to-peer transmission of tobacco harms is an effective strategy in reducing use because peers are less intimidating than physicians [@secker2000; @williams2011]. Most participants reported that they shared information from the presentations with others, although women in the RHP condition were not more likely to share than women in the GHP condition, contrary to predictions (see Figure S6 and Table 2, Share presentation model). This finding nonetheless indicates that sharing in this community is an effective strategy for communicating public health knowledge, and future research should consider using peer-to-peer transmission of tobacco harms to promote cessation in this community.

One exploratory result was particularly noteworthy. The women who were arguably most heavily influenced by the presentations where those who exhibited the greatest increases in the number of free-listed harms at followup relative to baseline. In support, an exploratory analysis found that an increase in the freelisted number of reproductive harms, but not general harms, was a significant negative predictor of cotinine. See Table S1 and Figure S5.

A second exploratory analysis found that probable use of tobacco was positively associated with older age, income, and number of children, and a few other variables, and negatively associated with being married and pregnant. See Figure S4. The positive effect of age is consistent with the higher prevalence of tobacco use seen globally in older women [@ng2014], and this pattern is consistent with the fetal protection hypothesis because the older women in our study were approaching the end of their reproductive careers [and therefore would be less likely to be, or become, pregnant; @hagen2013; @hagen2016]. The positive effect of income with use is consistent with gender inequality models [@hitchman2011], but the positive effect of number of children is not. The negative effect of married status with tobacco use is also seen in other populations [@ramsey2019], and the negative effect of pregnancy status is consistent with the fetal protection model.

@@

$$
penalty = (1-\alpha)||\beta||_2^2/2+\alpha||\beta||_1.
$$

@@

Individuals who exhibited the greatest increases in number of free-listed harms at followup were arguably most heavily influenced by the presentations. We therefore fit an exploratory mixed effects linear regression model of log10 followup cotinine as a function of the changes in numbers of free-listed general harms and reproductive harms, controlling for log10 baseline cotinine and Presentation. Increases in the numbers of free-listed reproductive harms, but not general harms, were negatively associated with followup cotinine. See Table S1.

@@

Table S1: Linear mixed effects model of log10 followup cotinine concentration as a function of the changes in numbers of general and reproductive harms free-listed at followup compared to baseline, controlling for log10 baseline cotinine and presentation condition. Values are estimated coefficients (standard errors).

```{r cotharmfig, fig.width=10, fig.cap="Effects plot for a regression model of log10 cotinine concentration versus the change in number of (A) general harms and (B) reproductive harms mentioned at followup compared to baseline (see Table S1)."}
p_m_cot_deltargharmsGH + p_m_cot_deltargharmsRH + plot_annotation(tag_levels='A')
```

@@

This table displays the coefficients of the same regression models as Table 2, but adds z-values, full p-values, and 95% CI's.

@@

Table S2. Regression coefficients, standard errors, z-values, p-values, and 95% confidence intervals for the models described in the main text and displayed in Table 2.
